Moral Foundations of U.S. Political News Organizations

### 2019-04-24 pilot_manuscript.Rmd
Another update; copy editing is next

Thus, the researchers' attention turned to the instrument used to detect differences: the Moral Foundations Dictionary. As mentioned before, MFD is a measure developed from Moral Foundations Theory [@Graham2009]. Based upon the results obtained, it might be necessary to investigate alternative instruments that could better elucidate the differences of interest. Likewise, other theoretical perspectives may be more equipped to explain political differences in discourse.

Tools such as these could be developed through the Linguistic Inquiry and Word Count program (LIWC) [@Pennebaker2007]. As stated before, the Moral Foundations Dictionary was established through an analysis of Christian sermons performed mostly with LIWC as well as some human-enforced corrections due to context [@Graham2009]. This approach was grounded in Moral Foundations Theory, a clear framework within which the authors developed the MFD. The original authors of the MFD established sound psychometric properties for the tool, implying its validity in several applications. For this reason, the researchers elected to utilize the MFD as their primary instrument.

Other instruments could theoretically be developed through a LIWC analysis, including those that would be useful for analyzing political discourse. Researchers could obtain percent occurrence scores for words appearing in certain LIWC categories, which could then be weighted with valence scores from @Warriner2013. The researchers will consider other applications of the LIWC in future studies aiming to analyze political content. @Pennebaker2007 has suggested the program could be applied to analyses of cognitive, social, and emotional states. Furthermore, there is some evidence LIWC can be used to detect deception, but the reliability of the program's accuracy in this regard (67%) calls this assertion into question [@Pennebaker2011]. Despite the promising extent of LIWC's abilities, without a firm theoretical foundation for such analyses, any such applications of LIWC would face serious questions as to their relevance, validity, and generalizability. Nevertheless, LIWC analyses may represent a path to further research into political differences in news media as well as theory formation.

@@

MAYBE DO DATA, THEN THEORIZE

>>>>

Thus, the researchers' attention turned to the instrument used to detect differences: the Moral Foundations Dictionary. As mentioned before, MFD is a measure developed from Moral Foundations Theory [@Graham2009]. Based upon the results obtained, it might be necessary to investigate alternative instruments that could better elucidate the differences of interest. Going beyond specific instruments, other theoretical perspectives may be more equipped to explain political differences in discourse. Likewise, an atheoretical approach in which large quantities of data are collected from which theories are formulated may be best suited to this area of research.

Alternative instruments could be developed through the Linguistic Inquiry and Word Count program (LIWC) [@Pennebaker2007]. As stated before, the Moral Foundations Dictionary was established through an analysis of Christian sermons performed mostly with LIWC as well as some human-enforced corrections due to context [@Graham2009]. This approach was grounded in Moral Foundations Theory, a clear framework within which the authors developed the MFD. The original authors of the MFD established sound psychometric properties for the tool, implying its validity in several applications. For this reason, the researchers elected to utilize the MFD as their primary instrument.

Instruments that improve upon the Moral Foundations Dictionary could theoretically be developed through a LIWC analysis. Researchers could obtain percent occurrence scores for words appearing in certain LIWC categories, which could then be weighted with valence scores from @Warriner2013. The researchers will consider other applications of the LIWC in future studies aiming to analyze political content. @Pennebaker2007 has suggested the program could be applied to analyses of cognitive, social, and emotional states. Furthermore, there is some evidence LIWC can be used to detect deception, but the reliability of the program's accuracy in this regard (67%) calls this assertion into question [@Pennebaker2011]. Despite the promising extent of LIWC's abilities, without a firm theoretical foundation for such analyses, any such applications of LIWC would face serious questions as to their relevance, validity, and generalizability. Nevertheless, LIWC analyses may represent a path to further research into political differences in news media as well as theory formation.

@@

Finally, future studies aiming to uncover political lean in discourse may benefit from a bottom-up, data-driven approach. Researchers might derive substantive insights into political lean through gathering data after which they may formulate theories that explain systematic observations obtained from that data. The availability of text corpora along with methods for extracting large amounts of text from the internet (as was demonstrated in the current study) potentially make this a feasible option. Likewise, there are several approaches to analyzing such data, including linear models (like multi-level models) and network-style models such as latent semantic analysis [@Landauer1998]. Owing to the wealth of representative data as well as the sophistication of current analytic tools, there is high potential for the explanatory power of new theories involving political discourse.
